---
title: 'Project 7 Task 1: Store Format'
output:
  html_document:
    df_print: paged
---

# Description
Determine the optimal number of store formats based on sales data.
- Sum sales data by StoreID and Year
- Use percentage sales per category per store for clustering (category sales
as a percentage of total store sales)
- Use only 2015 sales data
- Use a K-means clustering model

# Task 1: Store Format
## Import Data
```{r}
demographic <- read.csv('data/storedemographicdata.csv')
info <- read.csv('data/storeinformation.csv')
sales <- read.csv('data/storesalesdata.csv')
```

## Data Wrangling
```{r}
library(dplyr)

# Sum sales categories by store and year
sum_sales <- sales %>% group_by(Store, Year) %>% 
  summarise(dry_grocery=sum(Dry_Grocery),
            dairy=sum(Dairy),
            frozen_food=sum(Frozen_Food),
            meat=sum(Meat),
            produce=sum(Produce),
            floral=sum(Floral),
            deli=sum(Deli),
            bakery=sum(Bakery),
            general_merchandise=sum(General_Merchandise))
# Sum of all categories
sum_sales$total <- sum_sales$dry_grocery + sum_sales$dairy + sum_sales$frozen_food + sum_sales$meat + sum_sales$produce + sum_sales$floral + sum_sales$deli + sum_sales$bakery + sum_sales$general_merchandise

# Determine percentage of total
sum_sales_percentage <- data.frame(store=sum_sales$Store,
                                   year=sum_sales$Year,
                                   dry_grocery=100*sum_sales$dry_grocery/sum_sales$total,
                                   dairy=100*sum_sales$dairy/sum_sales$total,
                                   frozen_food=100*sum_sales$frozen_food/sum_sales$total,
                                   meat=100*sum_sales$meat/sum_sales$total,
                                   produce=100*sum_sales$produce/sum_sales$total,
                                   floral=100*sum_sales$floral/sum_sales$total,
                                   deli=100*sum_sales$deli/sum_sales$total,
                                   bakery=100*sum_sales$bakery/sum_sales$total,
                                   general_merchandise=100*sum_sales$general_merchandise/sum_sales$total)

# Check to make sure each column adds up to 100
sum_sales_percentage$total <- sum_sales_percentage$dry_grocery + sum_sales_percentage$dairy + sum_sales_percentage$frozen_food + sum_sales_percentage$meat + sum_sales_percentage$produce + sum_sales_percentage$floral + sum_sales_percentage$deli + sum_sales_percentage$bakery + sum_sales_percentage$general_merchandise

# Obtain 2015 sales data
sum_sales_percentage_2015 <- subset(sum_sales_percentage, year==2015)
```

## Cluster Validation
```{r}
# Load dependencies
library(NbClust)

# Determine optimal number of clusters
NbClust(sum_sales_percentage_2015[3:11], method='kmeans', index='ch')
summary <- NbClust(sum_sales_percentage_2015[3:11], method='kmeans', index='all')
```
Per the Calinski-Harabasz Method, as well as 4 other cluster validation methods,
3 clusters should be used.

## Build Cluster Model
```{r}
# Load dependencies
library(robustHD)

# Build Model around standardized data
fit <- kmeans(standardize(sum_sales_percentage_2015[3:11]), 3)

# Attach clusters to dataframe
sum_sales_percentage_2015$cluster <- fit$cluster
sum_sales_percentage_2015$total_sales <- subset(sum_sales, Year==2015)$total

# Export data
write.csv(sum_sales_percentage_2015, 'data/sum_sales_percentage_2015_clusters.csv')
```

## Conclusions

1. What is the optimal number of store formats? How did you arrive at that 
number?
The optimal number of store formats is 3. I determined this by first summarizing 
sales data for each store in 2015. Then, I determined the optimal number of 
clusters assuming a kmeans cluster method through the Calinski-Harabasz, 
and numerous other, cluster validation methods. 

2. How many stores fall into each store format?
Per the table below, cluster 1 contains 17 stores, cluster 2 contains 35 stores,
and cluster 3 contains 33 stores.
```{r echo=FALSE}
sum_sales_percentage_2015 %>% group_by(cluster) %>% summarise(count=n())
```

3. Based on the results of the clustering model, what is one way that the 
clusters differ from one another?
The table below summarizes the differences between the 3 clusters. One way that
the clusters differ from each other, is the percentage of general merchandise
sales, with Cluster 1 at the highest percentage and Cluster 2 at the lowest
percentage.
```{r echo=FALSE}
sum_sales_percentage_2015 %>% group_by(cluster)

sum_sales_percentage_2015 %>% group_by(cluster) %>% 
  summarise(dry_grocery=mean(dry_grocery),
            dairy=mean(dairy),
            frozen_food=mean(frozen_food),
            meat=mean(meat),
            produce=mean(produce),
            floral=mean(floral),
            deli=mean(deli),
            bakery=mean(bakery),
            general_merchandise=mean(general_merchandise))
```

4. Please provide a map created in Tableau that shows the location of the 
existing stores, uses color to show cluster, and size to show total sales. Make 
sure to include a legend! Feel free to simply copy and paste the map into the 
submission template.
The below Tableau map displays the store locations, colored by cluster.
https://public.tableau.com/profile/michael.gysel#!/vizhome/StoreClusters_10/Dashboard1?publish=yes 






