---
title: 'Project 7 Task 2: Determine the Store Format for New Stores'
output:
  html_document:
    df_print: paged
---

# Description
The grocery store chain has 10 new stores opening up at the beginning of the 
year. The company wants to determine which store format each of the new stores 
should have. However, we don’t have sales data for these new stores yet, so 
we’ll have to determine the format using each of the new store’s demographic 
data.


# Classification Models
1. Decision Tree
2. Random Forest
3. Boosted Model


# Data Wrangling
## Gather
```{r}
demographic <- read.csv('data/storedemographicdata.csv')
info <- read.csv('data/storeinformation.csv')
sales <- read.csv('data/storesalesdata.csv')
clusters <- read.csv('data/sum_sales_percentage_2015_clusters.csv')
```

## Clean
```{r}
# Add demographic to info
demo_cluster <- demographic[0:85,2:45]
demo_cluster$cluster <- clusters$cluster
```

## Train and Test Datasets
```{r}
library(dplyr)

set.seed(3)
# Train (80% of observations)
demo_train <- sample_n(demo_cluster, round(0.8*nrow(info)))
# Test (20% of observations)
demo_test <- demo_cluster[!(rownames(demo_cluster) %in% rownames(demo_train)),]
```


# Build Models
## 1. Decision Tree
### Build Decision Tree
```{r}
# Import libraries
library(rpart)

# Grow tree
mytree <- rpart(cluster ~ ., data=demo_train)

# Display the results
printcp(mytree)
summary(mytree)
```

### Score Model
```{r}
# Load dependency
library('caret')

# Test Data
predict <- data.frame('Prediction'=round(predict(mytree, demo_test)))
predict
```

### Confusion Matrix
```{r}
# Confusion matrix - Yes/no are actual. False/True are predicted
cm.DecisionTree <- confusionMatrix(factor(demo_test$cluster, levels=c(1,2,3)), factor(predict$Prediction, levels=c(1,2,3)))
cm.DecisionTree
```
The Decision Tree Model has an accuracy of 0.667.

## 2. Random Forest Model
### Build Random Forest
```{r}
# Load dependencies
library('randomForest')

# do.trace displays the error rate vs. sample size
fit <- randomForest(cluster ~ ., data=demo_train, importance=TRUE, ntree=100, do.trace=FALSE)
```

### Score the Model
```{r}
# Predict test dataset
predict <- data.frame('Prediction'=round(predict(fit, demo_test, type='class')))
predict
```

### Confusion Matrix
```{r}
# Confusion matrix - Yes/no are actual. False/True are predicted
cm.RandomForest <- confusionMatrix(factor(demo_test$cluster, levels=c(1,2,3)), factor(predict$Prediction, levels=c(1,2,3)))
cm.RandomForest
```
The Random Forest Model has an accuracy of 0.778.

## 3. Boosted Model
### Build Boosted Model
```{r}
# Dependencies
library('gbm')

# Boosted Model
boost_model <- gbm(cluster ~ ., data=demo_train, n.trees=1580)

# Check the best number of iterations
best.iter = gbm.perf(boost_model)
best.iter
```

### Score the Model
```{r}
# Score the model
predict <- data.frame('Prediction'=round(predict.gbm(boost_model, newdata=demo_test, n.trees=1580, type='response')))
predict
```

### Confusion Matrix
```{r}
# Make sure columns are factor type with 3 levels
demo_test$cluster <- factor(demo_test$cluster, levels=c(1,2,3))
predict$Prediction <- factor(predict$Prediction, levels=c(1,2,3))

# Confusion matrix
confusionMatrix(demo_test$cluster, predict$Prediction)
```
The Boosted Model has an accuracy of 0.667.

# Predict Store Format
Use the Random Forest Model to predict the best store format for each of the 10 
new stores.

## Build Full Random Forest Model
```{r}
# Random Forest Model
rf_model <- randomForest(cluster ~ ., data=demo_cluster, importance=TRUE, ntree=100, do.trace=FALSE)
```

## Score the 10 New Stores
```{r}
# Score the 10 New Stores
demo_new <- demographic[86:95,2:45]
predict <- data.frame('Prediction'=round(predict.gbm(boost_model, newdata=demo_new, n.trees=1776, type='response')))

# Add to demographic dataframe
info$cluster <- c(demo_cluster$cluster, predict$Prediction)

# Export data
write.csv(info, 'data/info_cluster.csv')
```


# Conclusions

1. What methodology did you use to predict the best store format for the new 
stores? Why did you choose that methodology?
I used the Random Forest Model because this had the greatest accuracy when 
predicting correct clusters of the 85 existing stores.

2. What are the three most important variables that help explain the 
relationship between demographic indicators and store formats? Please include a 
visualization.
Per the variable importance plot of the boosted model below, the three most
important factors are percentage of population ages 0 to 9, percentage of 
population with a Bachelor's Degree, and home value of atleast $750k.
```{r echo=FALSE}
# Variable Importance Factors
varImpPlot(rf_model)
```

3. What format do each of the 10 new stores fall into? Please provide a data 
table.
The data table below shows the 10 new stores and corresponding clusters.
```{r echo=FALSE}
info[85:95,c(1,7)]
```