---
title: 'Project 7 Task 3: Forecasting'
output:
  html_document:
    df_print: paged
---

# Description
You’ve been asked to prepare a monthly forecast for produce sales for the full 
year of 2016 for both existing and new stores. To do so, follow the steps below.
- Forecast produce sales for existing stores
- Forecast produce sales for new stores


# Data Wrangling
## Gather
```{r}
# Import Data
demographic <- read.csv('data/storedemographicdata.csv')
info <- read.csv('data/info_cluster.csv')
sales <- read.csv('data/storesalesdata.csv')
```
## Clean
```{r}
# Merge sales data with clusters
avg_sales_total <- merge(sales, info, by.x='Store', by.y='Store')
```


# Time Series Forecasting
## Train and Test Datasets
```{r}
library(dplyr)

# Aggregate produce sales data across all stores in each cluster by month
avg_sales <- avg_sales_total %>% group_by(Store, Year, Month) %>% 
  summarise(avg_produce=sum(Produce)) %>%
  group_by(Year, Month) %>%
  summarise(avg_produce=mean(avg_produce))

# Train (to 6/15)
avg_sales_train <- head(avg_sales, 40)
# Test (7/15 to 12/15)
avg_sales_test <- tail(avg_sales, 6)
```
## Convert Data to Time Series
```{r}
# Load dependencies
library(PerformanceAnalytics)

# Convert avg_sales to time series object
ts_train <- ts(avg_sales_train$avg_produce, start=c(2012, 3), end=c(2015, 6), frequency=12)
ts_full <- ts(avg_sales$avg_produce, start=c(2012, 3), end=c(2015, 12), frequency=12)
```


## ETS Model
### Determine ETS Components
```{r}
# Fit time series decomposition
fit <- stl(ts_train, s.window='period')
# Plot
plot(fit)
```
As the above time series decomposition plot shows, the time series displays:

- Error: Multiplicative
- Trend: None
- Seasonality: Multiplicative

### Build ETS Model
```{r}
# Load dependencies
library(forecast)

# ETS Model with train dataset
fit_ets <- ets(ts_train, model='MNM')
plot(forecast(fit_ets))
```

### Forecast/Test Accuracy
```{r}
# fit_ets is the model prediction
# ts_full is the actual time series object
accuracy(forecast(fit_ets, 6), ts_full[41:46])

# Plot accuracy
# Fitted in Red
plot(fit_ets$fitted, col='red')
# Actual in Blue
lines(ts_full, col='blue')
```
The ets model yields an MAPE of 1.92%.

## ARIMA Model
### Stationarize Dataset
Plot the data to check if stationary
```{r}
# Plot data to check if constant mean/variance
plot(ts_train)
```
The data is not stationary and is seasonal, so let's seasonally stationarize
the data.
```{r}
# First Seasonal Difference
avg_sales$first_difference <- c(rep(NA,12), diff(avg_sales$avg_produce, lag=12))
```
Plot the data again, to check if stationary
```{r}
# Make first_difference time series
ts_fd <- ts(avg_sales$first_difference[1:40], start=c(2012, 3), end=c(2015, 6), frequency=12)

# Plot first_difference
plot(ts_fd)
```
The first_difference does not appear seasonal, but is still not stationary. 
Let's take a second, non-seasonal difference.
```{r}
# Second, non-seasonal difference
avg_sales$second_difference <- c(NA, diff(avg_sales$first_difference, lag=1))
```
Plot the data again, to check if stationary
```{r}
# Make second_difference time series
ts_sd <- ts(avg_sales$second_difference[1:40], start=c(2012, 3), end=c(2015, 6), frequency=12)

# Plot first_difference
plot(ts_sd)
```
Now, the time series displays a constant mean and variance, without any 
seasonality.

The model structure thus far, after taking a seasonal (D=1) difference and 
non-seasonal difference (d=1) to stationarize the data, with a period of 12 is:
- ARIMA(0,1,0)(0,1,0)[12]

### AR and MA Terms
#### ACF Plot
```{r}
# Plot the ACF of second_difference
ggAcf(avg_sales$second_difference[1:40], lag.max=48)
```
#### PACF Plot
```{r}
# Plot the PACF of second_difference
ggPacf(avg_sales$second_difference[1:40], lag.max=48)
```
- The ACF and PACF have negative values at lag 1, suggesting a non-seasonal MA
Term, signified as q=1. 
- The ACF and PACF have negative values at lag 12, suggesting a seasonal MA
Term, signified as Q=1.

Thus, the model structure is:
- ARIMA(0,1,1)(0,1,1)[12]

### Build Model
```{r}
# Build Auto Arima Model
fit_arima <- Arima(ts_train, order=c(0,1,1), seasonal=list(order=c(0,1,1), period=12))
fit_arima
```

### Forecast/Test Accuracy
```{r}
# fit_arima is the model prediction
# ts_full is the actual time series object
accuracy(forecast(fit_arima, 6), ts_full[41:46])

# Plot accuracy
# Fitted in Blue
plot(forecast(fit_arima))
# Actual in Red
lines(ts_full, col='red')
```
The arima model yields an MAPE of 2.82%.

## Predict Next 12 Periods

### Choose Best Model
The ets model yields the  highest accuracy.

### Build Model with All Data
Build the ETS Model with all data
```{r}
# Arima Model with all data
fit_bm <- fit_ets <- ets(ts_full, model='MNM')
plot(forecast(fit_bm))
```

### Forecast Next 12 Periods
```{r}
# ARIMA Model with all data
forecast <- forecast(fit_bm, 12)
forecast
```


# Conclusions

1. What type of ETS or ARIMA model did you use for each forecast? 
Use ETS(a,m,n) or ARIMA(ar, i, ma) notation. How did you come to that decision?
An ETS(M,N,M) Model was used. After assessing the time series decomposition
plot, it became clear the Error is Multiplicative, Trend is None, and 
Seasonality is Multiplicative. This ETS Model yielded a higher accuracy than the
ARIMA Model, so it was chosen for the time series forecast.

2. Please provide a table of your forecasts for existing and new stores. Also, 
provide visualization of your forecasts that includes historical data, existing 
stores forecasts, and new stores forecasts.
The table below shows the produce sales forecasts for existing and new stores 
for 2016.
The link below shows the data visualization of total produce sales.
https://public.tableau.com/profile/michael.gysel#!/vizhome/P7-TotalProduceSales/Dashboard1?publish=yes 
```{r echo=FALSE}
# Build forecast tables
forecasts <- data.frame('Existing Stores' = forecast$mean*85,
                        'New Stores' = forecast$mean*10)

# Change rownames
rownames(forecasts) <- c('Jan-16', 'Feb-16', 'Mar-16', 'Apr-16', 'May-16', 'Jun-16', 'Jul-16', 'Aug-16', 'Sep-16', 'Oct-16', 'Nov-16', 'Dec-16')

# Display table
forecasts
```

```{r echo=FALSE}
# Export data for data visualization
# Total monthly sales

# Forecast
sales_forecast <- data.frame('Year'=as.integer(rep(c(2016),12)),
                             'Month'=as.integer(seq(1,12,1)),
                             'Existing'=as.numeric(forecasts$Existing.Stores),
                             'New'=as.numeric(forecasts$New.Stores))

# Past sales
sales_past <- avg_sales_total %>% group_by(Year, Month) %>% 
  summarise(avg_produce=sum(Produce))

# Timeframe
dates <- seq(as.Date("2012/3/1"), as.Date("2016/12/1"), by = "month")
# Existing
existing_data <- c(sales_past$avg_produce, sales_forecast$Existing)
# New
new_data <- c(rep(NA,46),sales_forecast$New)
# Combine for all sales data
sales_forecasts <- data.frame('Date'=dates,
                              'Existing'=existing_data, 
                              'New'=new_data)

# Export data
write.csv(sales_forecasts, 'data/sales_forecasts.csv')
```




