---
title: 'Project 7: Conclusions'
output:
  pdf_document: default
  html_notebook: default
---

# Task 1
1. What is the optimal number of store formats? How did you arrive at that 
number?
The optimal number of store formats is 3. I determined this by first summarizing 
sales data for each store in 2015. Then, I determined the optimal number of 
clusters assuming a kmeans cluster method through the Calinski-Harabasz, 
and numerous other, cluster validation methods. 

2. How many stores fall into each store format?
Per the table below, cluster 1 contains 17 stores, cluster 2 contains 35 stores,
and cluster 3 contains 33 stores.
```{r echo=FALSE}
sum_sales_percentage_2015 %>% group_by(cluster) %>% summarise(count=n())
```

3. Based on the results of the clustering model, what is one way that the 
clusters differ from one another?
The table below summarizes the differences between the 3 clusters. One way that
the clusters differ from each other, is the percentage of general merchandise
sales, with Cluster 1 at the highest percentage and Cluster 2 at the lowest
percentage.
```{r echo=FALSE}
sum_sales_percentage_2015 %>% group_by(cluster)

sum_sales_percentage_2015 %>% group_by(cluster) %>% 
  summarise(dry_grocery=mean(dry_grocery),
            dairy=mean(dairy),
            frozen_food=mean(frozen_food),
            meat=mean(meat),
            produce=mean(produce),
            floral=mean(floral),
            deli=mean(deli),
            bakery=mean(bakery),
            general_merchandise=mean(general_merchandise))
```

4. Please provide a map created in Tableau that shows the location of the 
existing stores, uses color to show cluster, and size to show total sales. Make 
sure to include a legend! Feel free to simply copy and paste the map into the 
submission template.
The below Tableau map displays the store locations, colored by cluster.
https://public.tableau.com/profile/michael.gysel#!/vizhome/StoreClusters_10/Dashboard1?publish=yes 


# Task 2

1. What methodology did you use to predict the best store format for the new 
stores? Why did you choose that methodology?
I used the Random Forest Model because this had the greatest accuracy when 
predicting correct clusters of the 85 existing stores.

2. What are the three most important variables that help explain the 
relationship between demographic indicators and store formats? Please include a 
visualization.
Per the variable importance plot of the boosted model below, the three most
important factors are percentage of population ages 0 to 9, percentage of 
population with a Bachelor's Degree, and home value of atleast $750k.
```{r echo=FALSE}
# Variable Importance Factors
varImpPlot(rf_model)
```

3. What format do each of the 10 new stores fall into? Please provide a data 
table.
The data table below shows the 10 new stores and corresponding clusters.
```{r echo=FALSE}
info[85:95,c(1,7)]
```


# Task 3

1. What type of ETS or ARIMA model did you use for each forecast? 
Use ETS(a,m,n) or ARIMA(ar, i, ma) notation. How did you come to that decision?
An ETS(M,N,M) Model was used. After assessing the time series decomposition
plot, it became clear the Error is Multiplicative, Trend is None, and 
Seasonality is Multiplicative. This ETS Model yielded a higher accuracy than the
ARIMA Model, so it was chosen for the time series forecast.

2. Please provide a table of your forecasts for existing and new stores. Also, 
provide visualization of your forecasts that includes historical data, existing 
stores forecasts, and new stores forecasts.
The table below shows the produce sales forecasts for existing and new stores 
for 2016.
The link below shows the data visualization of total produce sales.
https://public.tableau.com/profile/michael.gysel#!/vizhome/P7-TotalProduceSales/Dashboard1?publish=yes 
```{r echo=FALSE}
# Build forecast tables
forecasts <- data.frame('Existing Stores' = forecast$mean*85,
                        'New Stores' = forecast$mean*10)

# Change rownames
rownames(forecasts) <- c('Jan-16', 'Feb-16', 'Mar-16', 'Apr-16', 'May-16', 'Jun-16', 'Jul-16', 'Aug-16', 'Sep-16', 'Oct-16', 'Nov-16', 'Dec-16')

# Display table
forecasts
```



